---
title: "Car Specification Impact on Miles Per Galleon (MPG) Performance"
author: "Jason Lim"
date: "Friday, June 05, 2015"
output: pdf_document
---

#Executive Summary
The relationship between MPG and a set of variables is analyzed utilizing a data set extracted from an Automobile Magazine with emphasis on the relationship between MPG and the type of transmission, Auto or Manual.

The average MPG for an automatic car is approximately **17 MPG** whereas for manual car is approximately **22 MPG**. 

After the data has been analysed using the Linear Regression Model, it is also determined that with **all other variables set to a constant**, manual cars' performance in MPG field is approximately **1.8 MPG** better compared to automatic cars.

#Exploratory Analysis

The dataset is loaded and explored.
```{r}
data(mtcars)
```

The variables **cyl**, **vs**, **gear**, **carb**, and **am** are originally classified as num, but they're actually factors, thus some cleaning up was required. Additionally,  **am**'s 0 and 1 values has been replaced by auto and manual respectively.

```{r}
mtcars$cyl  <- factor(mtcars$cyl)
mtcars$vs   <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am   <- factor(mtcars$am,labels=c("Auto","Manual"))
```

Now exploratory analysis can be done. As the relationship between MGP and Auto/Manual type cars is the focus, a boxplot featuring these 2 is plotted as seen in **Figure 1**.

From **Figure 1**, it is clear that on average a manual car have a performance of approximately **22 MPG** whereas an automatic car has a performance of approximately **17 MPG**.

Deeper analysis is done using the Linear Regression Method.

#Analysis
Firstly, all variables are taken into account when predicting MPG before the variables that are most impactful is narrowed down.

```{r}
allModel<-lm(mpg ~ .,data=mtcars)
summary(allModel)$adj.r.squared
```
It is not practical to have so many predictors in a Model, and the R-Squared value is just **0.78**, meaning the model covers only 78% of variability. Thus it is recommended that a different model be used.

The Step function will be used to further improve the above model :
```{r,include=F}
stepModel<-step(allModel,direction="both")
```

```{r}
stepModel$coef
summary(stepModel)$adj.r.squared
```

From the above, most of the predictors have been eliminated, leaving just **cyl**, **wt**, **am** as main predictors, the adjusted R-Squared value has also been improved to **0.84**. 

#ANOVA Test
An **ANOVA** test comparing **stepModel**, and the base model consisting only **am** as predictor is performed.

```{r}
baseModel<-lm(mpg ~ am,data=mtcars)
anova(baseModel,stepModel)[6]
```
As can be seen from the P value of the test, the **null hypothesis** that states cyl, hp, wt does not contribute is rejected.

#Residuals vs Fitted
**Figure 2** shows the **Residuals vs Fitted** plot, and it shows that **Toyota Corolla**, **Datsun 710**, and **Fiat 128** are the main residuals that has caused the curve to occur.

#Conclusion
Manual cars have **higher MPG** compared to Automatic cars by 1.8 MPG. Car Weight's affect on car weight is the highest, with MPG decreasing at an approximate rate of **2.5MPG per 1000lb increase in weight**.

#Appendix
Figure 1 - Boxplot of MGP and A/M
```{r}
boxplot(mpg ~ am, data = mtcars, ylab = "Miles Per Gallon", xlab = "Transmission Type")
```

Figure 2 - Residual vs Fitted
```{r}
plot(stepModel,which=c(1,1))
```

